---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ymd

<!-- badges: start -->
[![R-CMD-check](https://github.com/shrektan/ymd/workflows/R-CMD-check/badge.svg)](https://github.com/shrektan/ymd/actions)
<!-- badges: end -->

Handle common ymd Date Operations in R using Rust. It converts ymd integers or strings to Date, e.g., `211225` to `as.Date("2021-12-25")` and it provides addition helper functions like bop or eop (quick finding the begining or ending of period, e.g., the 1st date of the year or month).

It's similar to the `lubridate` package but will be much lighter and focuses on Date objects.

## Installation

Note, in order to compile this package from source you need the Rust toolchain, which can be found in [Rust's official website](https://www.rust-lang.org).

## Some use case and benchmarks

```{r bench}
x <- c("210101", "21/02/03", "89-1-03", "1989.03.05")
x <- rep(x, 100)
bench::mark(
  ymd::ymd(x),
  lubridate::ymd(x), time_unit = "us"
)

x <- c(210101, 210224, 211231, 19890103)
x <- rep(x, 100)
bench::mark(
  ymd::ymd(x),
  lubridate::ymd(x), time_unit = "us"
)

x <- c("2021-01-01", "2022-12-31", "1995-03-22")
x <- rep(x, 100)
bench::mark(
  ymd::ymd(x),
  lubridate::ymd(x), time_unit = "us",
  as.Date(x)
)

x <- ymd::ymd(210515) + 1:100
bench::mark(
  ymd::eop$tm(x),
  lubridate::ceiling_date(x, "month") - 1
)

`%m+%` <- lubridate::`%m+%`
x <- ymd::ymd(c(200115, 200131, 200229, 200331, 200401))
bench::mark(
  ymd::edate(x, 2),
  x %m+% months(2)
)
bench::mark(
  ymd::edate(x, -12),
  x %m+% months(-12)
)
```
